---
title: "p8105_hw3_sdc2157"
author: "Stephanie Calluori"
date: 2023-10-14
output: github_document
---

```{r, data import, results = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)

data("instacart")

data("brfss_smart2010")

demographic_raw <- read_csv("data/nhanes_covar.csv", skip = 4, col_names = TRUE)
head(demographic_raw)
tail(demographic_raw)
colnames(demographic_raw)

accel_raw <- read_csv("data/nhanes_accel.csv", col_names = TRUE)
head(accel_raw)
tail(accel_raw)
colnames(accel_raw)
```

```{r, setup, echo = FALSE, results = FALSE, message = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

```{r}
instacart_clean <- instacart |> 
  janitor::clean_names()

head(instacart_clean)

str(instacart_clean)

sum(is.na(instacart_clean))

num_users <- instacart_clean |> 
  distinct(user_id)

```


The data set describes

The data set contains `r nrow(instacart_clean)` observations and `r ncol(instacart_clean)`. Each observation corresponds to a distinct instacart product. Variables include `r colnames(instacart_clean)`.

```{r}
num_aisle <- instacart_clean |> 
  count(aisle_id) |> 
  nrow()

aisle_items <- instacart_clean |> 
  select(aisle, product_name) |> 
  group_by(aisle) |> 
  summarize(
    n_items = n()
  ) |> 
  arrange(desc(n_items))

```

fresh veggies, fresh fruits, and packaged vegtables/fruits

```{r}
plot <- aisle_items |> 
  filter(n_items > 10000) |> 
  ggplot(aes(x = aisle, y = n_items)) +
  geom_col()

```

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
pop_items <- instacart_clean |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |> 
  summarize(n_obs = n()) |> 
  arrange(desc(n_obs)) |> 
  mutate(rank = min_rank(desc(n_obs))) |>
  filter(rank < 4) |> 
  knitr::kable(digits = 2)
  
  
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

