---
title: "p8105_hw3_sdc2157"
author: "Stephanie Calluori"
date: 2023-10-14
output: github_document
---

# Load packages

```{r, load packages, message = FALSE}
library(tidyverse)
library(p8105.datasets)

```

# Data Import

```{r, data import, results = FALSE, message = FALSE}
data("instacart")

data("brfss_smart2010")

demographic_raw <- read_csv("data/nhanes_covar.csv", skip = 4, col_names = TRUE)

accel_raw <- read_csv("data/nhanes_accel.csv", col_names = TRUE)

```

```{r, setup, echo = FALSE, results = FALSE, message = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "right"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

# Problem 1

```{r, clean and summarize instacart}
instacart_clean <- instacart |> 
  janitor::clean_names()

na_num <- sum(is.na(instacart_clean))

popular_items <- instacart_clean |> 
  group_by(product_name) |> 
  summarize(product_n = n()) |> 
  arrange(desc(product_n))

```

The instacart data set contains `r nrow(instacart_clean)` observations and `r ncol(instacart_clean)` variables. Each observation corresponds to a product from an order. There are `r instacart_clean |> distinct(order_id) |> nrow()` orders from `r instacart_clean |> distinct(user_id) |> nrow()` users (i.e., one order per user). 

Variables include `r colnames(instacart_clean)`. All variables are integers except for 4 character variables. NA values are not present in the data set. 

Across the orders, a total of `r instacart_clean |> distinct(product_id) |> nrow()` distinct products were purchased. The most purchased item was `r popular_items[1,1]`. 

```{r, aisle summaries}
num_aisle <- instacart_clean |> 
  count(aisle_id) |> 
  nrow()

aisle_items <- instacart_clean |> 
  group_by(aisle) |> 
  summarize(n_items = n()) |> 
  arrange(desc(n_items))

```

Customers ordered from `r num_aisle` different aisles. The top 3 aisles from which the most items were ordered were `r aisle_items[1,1]`, `r aisle_items[2,1]`, and `r aisle_items[3,1]`.

```{r, aisle plot}
aisle_items |> 
  filter(n_items > 10000) |> 
  mutate(aisle = forcats::fct_reorder(aisle, n_items, .desc = TRUE)) |> 
  ggplot(aes(x = aisle, y = n_items)) +
  geom_col() +
  labs(
    title = "Number of items ordered from each aisle",
    x = "Aisle",
    y = "Number of items",
    caption = "*Only includes aisles from which >10,000 items were ordered"
  ) +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))

```

Following the top 3 aisles, yogurt, packaged cheese, and walter seltzer sparkling water were the most popular aisles.

```{r, aisle table, message = FALSE}
instacart_clean |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |> 
  summarize(times_ordered = n()) |> 
  arrange(desc(times_ordered)) |> 
  mutate(rank = min_rank(desc(times_ordered))) |>
  filter(rank < 4) |> 
  knitr::kable(digits = 2)
  
```

The top item ordered in the packaged vegetables fruits aisle was organic baby spinach. In the baking ingredients aisle, light brown sugar was the most ordered item. Snack sticks chicken & rice recipe dog treats was the top item ordered in the dog food care aisle.

The table below shows the mean hour of the day at which Coffee Ice Cream and Pink Lady Apples were ordered on each day of the week. Overall, these items were typically ordered between late morning and mid-afternoon, regardless of the day of the week.


```{r, order time table, message = FALSE}
instacart_clean |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  mutate(
    order_dow = recode(
      order_dow, 
      "0" = "Sunday",
      "1" = "Monday",
      "2" = "Tuesday",
      "3" = "Wednesday",
      "4" = "Thursday",
      "5" = "Friday",
      "6" = "Saturday"),
  ) |> 
  group_by(product_name, order_dow) |> 
  arrange(product_name) |> 
  summarize(avg_hr = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = avg_hr
  ) |>   
  select(product_name, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) |> 
  knitr::kable(digits = 2)

```


# Problem 2

```{r, cleaned brfss, results = FALSE}
brfss_clean <- brfss_smart2010 |> 
  janitor::clean_names() |> 
  rename(state = locationabbr, county = locationdesc) |> 
  filter(topic == "Overall Health", 
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |> 
  mutate(response = forcats::fct_relevel(response, 
                                c("Poor", "Fair", "Good", "Very good", "Excellent")))

state_num <- brfss_clean |> 
  distinct(state) |> 
  count()

unique(brfss_clean$state)

county_num <- brfss_clean |> 
  distinct(county) |> 
  count()

```

The cleaned brfss data set focusing on the topic of overall health contains `r nrow(brfss_clean)` observations and `r ncol(brfss_clean)` variables. These data were collected from `r county_num` counties across the 50 states and Washington, D.C. from `r min(pull(brfss_clean, year))` to `r max(pull(brfss_clean, year))`.

The table below lists the states where 7 or more counties were surveyed in 2002. Six states met these criteria. Pennsylvania was the highest with 10 counties surveyed.

```{r, states counties 2002 table}
brfss_clean |> 
  select(year, state, county) |> 
  filter(year == "2002") |> 
  group_by(state) |> 
  distinct(county) |>
  summarize(num_county = n()) |> 
  filter(num_county >= 7) |> 
  arrange(desc(num_county)) |> 
  knitr::kable(digits = 2)

```

The table below lists the states where 7 or more counties were surveyed in 2010. 14 states met these criteria. Florida was the highest with 41 counties surveyed.

```{r, states counties 2010 table}
brfss_clean |> 
  select(year, state, county) |> 
  filter(year == "2010") |> 
  group_by(state) |>
  distinct(county) |> 
  summarize(num_county = n()) |> 
  filter(num_county >= 7) |> 
  arrange(desc(num_county)) |> 
  knitr::kable(digits = 2)

```

```{r, excellent responses plot, message = FALSE}
brfss_clean |> 
  select(year, state, county, response, data_value) |>
  drop_na() |> 
  filter(response == "Excellent") |> 
  group_by(state, year) |> 
  summarize(avg_data_value = mean(data_value)) |> 
  ggplot(aes(x = year, y = avg_data_value, group = state, color = state)) +
  geom_line() +
  labs(
    title = "Avg value of excellent responses",
    caption = "*Responses relate to the Overall Health topic")
  
```

The figure above shows the average data value for excellent responses across the 50 states and Washington, D.C. from 2002 to 2010. The responses are in relation to the Overall Health topic. Some states show great variability in the average data value for excellent responses across this time period.

```{r, ny data values plot}
brfss_clean |> 
  select(year, state, response, data_value) |> 
  filter(year %in% c(2006, 2010), state == "NY") |> 
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(. ~ year) +
  labs(
    title = "Distributions of data values for responses in NY in 2006 & 2010",
    caption = "*Responses relate to the Overall Health topic") +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))

```

The figure above compares the distributions of data values for responses "Poor" to "Excellent" among counties in NY state during 2006 to those reported in 2010. The responses are in relation to the Overall Health topic. Notably, the median data value for "Very good" responses was greater in 2010 than 2006.


# Problem 3

```{r, cleaned mims data sets}
demographic_clean <- demographic_raw |> 
  janitor::clean_names() |> 
  mutate(sex = as.factor(sex),
         education = as.factor(education)
         ) |>
  mutate(sex = recode(sex, 
                      "1" = "male", 
                      "2" = "female"),
         education = recode(education, 
             "1" = "Less than high school",
             "2" = "High school equivalent",
             "3" = "More than high school")) |> 
  drop_na() |> 
  filter(age >= 21)

accel_clean <- accel_raw |> 
  janitor::clean_names()

combined_accel <- left_join(demographic_clean, accel_clean, by = "seqn")
  
```

Our combined data set contains `r nrow(combined_accel)` observations and `r ncol(combined_accel)` variables. The data set describes the demographics and the MIMS activity per minute over a 24 hr time period for each of the `r nrow(combined_accel)` participants.

The table below shows the number of males and females in each of the educational categories. Most males and females received more than a high school education.

```{r, education table, message = FALSE}
combined_accel |> 
  group_by(education, sex) |> 
  summarize(n_obs = n()) |> 
  pivot_wider(
    names_from = sex,
    values_from = n_obs,
  ) |> 
  knitr::kable(digits = 2)

```


```{r, education sex age plot}
combined_accel |> 
  ggplot(aes(x = sex, y = age)) +
  geom_boxplot() +
  scale_y_continuous(
    breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90),
    limits = c(0, 90)
  ) +
  labs(title = "Age distribution of males and females by education received",
       y = "Age (years)",
       x = "Sex") +
  facet_grid(. ~ education)

```

The figure above shows the age distribuion of males and females by education received. Among those who received less than a high school education, the median age of males and females was similar. Among those who received the equivalent of a high school education, the median age of females was greater than males. Among those who received more than a high school education, the median age of females was slightly less than that of males. 

```{r, total activity plot, message = FALSE}
combined_accel |>
  mutate(total_activity = rowSums(across(min1:min1440))) |> 
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    title = "Total MIMS activity across age for males and females",
    x = "Age (years)",
    y = "Total Activity (MIMS)",
    color = "Sex") +
  facet_grid(education ~ .) +
  theme(legend.position = "bottom") +
  theme(strip.text.y = element_text(angle = 0))

```

The figure above shows total MIMS activity across age for males and females. Among those who received the equivalent of a high school education and those who received more than a high school education, females generally had a higher total activity level than males across ages 21-80 years. In contrast, among those who received less than a high school education, females tended to have a higher total activity level until age 40; after age 40, males had a higher total activity level than females.

```{r, mims over time plot, message = FALSE}
combined_accel |> 
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    names_prefix = "min",
    values_to = "mims"
  ) |> 
  mutate(minute = as.numeric(minute)) |> 
  ggplot(aes(x = minute, y = mims, group = seqn, color = sex)) +
  geom_line() +
  geom_smooth(aes(group = sex)) +
  labs(
    title = "MIMS over a 24 hr period for males and females",
    x = "Minute",
    y = "MIMS",
    color = "Sex") +
  facet_grid(. ~ education) +
  theme(legend.position = "bottom")

```

The figure above shows the MIMS for males and females over a 24 hr period by education received. The 24 hr time period started at midnight. Generally, most people were active during the day. The MIMS activity for males and females who received more than a high school education displays a dramatic uptick at minute 500 and minute 8000.

