---
title: "p8105_hw3_sdc2157"
author: "Stephanie Calluori"
date: 2023-10-14
output: github_document
---

```{r, data import, results = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)

data("instacart")

data("brfss_smart2010")

demographic_raw <- read_csv("data/nhanes_covar.csv", skip = 4, col_names = TRUE)
head(demographic_raw)
tail(demographic_raw)
colnames(demographic_raw)

accel_raw <- read_csv("data/nhanes_accel.csv", col_names = TRUE)
head(accel_raw)
tail(accel_raw)
colnames(accel_raw)
```

```{r, setup, echo = FALSE, results = FALSE, message = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "right"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

```{r}
instacart_clean <- instacart |> 
  janitor::clean_names()

head(instacart_clean)

str(instacart_clean)

sum(is.na(instacart_clean))

num_users <- instacart_clean |> 
  distinct(user_id)

```


The data set describes

The data set contains `r nrow(instacart_clean)` observations and `r ncol(instacart_clean)`. Each observation corresponds to a distinct instacart product. Variables include `r colnames(instacart_clean)`.

```{r}
num_aisle <- instacart_clean |> 
  count(aisle_id) |> 
  nrow()

aisle_items <- instacart_clean |> 
  select(aisle, product_name) |> 
  group_by(aisle) |> 
  summarize(
    n_items = n()
  ) |> 
  arrange(desc(n_items))

```

fresh veggies, fresh fruits, and packaged vegtables/fruits

```{r}
plot <- aisle_items |> 
  filter(n_items > 10000) |> 
  ggplot(aes(x = aisle, y = n_items)) +
  geom_col()

```

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
pop_items <- instacart_clean |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |> 
  summarize(n_obs = n()) |> 
  arrange(desc(n_obs)) |> 
  mutate(rank = min_rank(desc(n_obs))) |>
  filter(rank < 4) |> 
  knitr::kable(digits = 2)
  
  
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.



```{r}
sub <- instacart_clean |> 
  drop_na() |> 
  select(product_name, order_dow, order_hour_of_day) |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  mutate(
    order_dow = recode(
      order_dow, 
      "0" = "Sunday",
      "1" = "Monday",
      "2" = "Tuesday",
      "3" = "Wednesday",
      "4" = "Thursday",
      "5" = "Friday",
      "6" = "Saturday"),
  ) |> 
  group_by(product_name, order_dow) |> 
  arrange(product_name) |> 
  summarize(avg_hr = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = avg_hr
  )


```


# Problem 2
50 states and Washington DC
```{r}
brfss_clean <- brfss_smart2010 |> 
  janitor::clean_names() |> 
  rename(state = locationabbr, county = locationdesc) |> 
  filter(topic == "Overall Health", 
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |> 
  mutate(response = fct_relevel(response, 
                                c("Poor", "Fair", "Good", "Very good", "Excellent")))

unique(brfss_clean$state)

```

In 2002, which states were observed at 7 or more locations? 

```{r}
brfss_clean |> 
  select(year, state, county) |> 
  filter(year == "2002") |> 
  group_by(state) |> 
  distinct(county) |>
  summarize(num_county = n()) |> 
  filter(num_county >= 7) |> 
  arrange(desc(num_county)) |> 
  knitr::kable(digits = 2)

```


In 2010, which states were observed at 7 or more locations?

```{r}
brfss_clean |> 
  select(year, state, county) |> 
  filter(year == "2010") |> 
  group_by(state) |>
  distinct(county) |> 
  summarize(num_county = n()) |> 
  filter(num_county >= 7) |> 
  arrange(desc(num_county)) |> 
  knitr::kable(digits = 2)

```

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

###Office Hrs ASK ABOUT: strategies to make more readable, check on group aesthetic?

```{r}
brfss_clean |> 
  select(year, state, county, response, data_value) |>
  drop_na() |> 
  filter(response == "Excellent") |> 
  group_by(state, year) |> 
  summarize(avg_data_value = mean(data_value)) |> 
  ggplot(aes(x = year, y = avg_data_value, color = state)) +
  geom_line() 
  
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_clean |> 
  select(year, state, response, data_value) |> 
  filter(year %in% c(2006, 2010), state == "NY") |> 
  mutate(response = fct_reorder(response, data_value)) |> 
  ggplot(aes(x = data_value, y = response)) +
  geom_density_ridges() +
  facet_grid(.~year)
```

#Problem 3
demogrpahic raw
accel raw

Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r}
demographic_clean <- demographic_raw |> 
  janitor::clean_names() |> 
  mutate(sex = as.factor(sex),
         education = as.factor(education)
         ) |>
  mutate(sex = recode(sex, 
                      "1" = "male", 
                      "2" = "female"),
         education = recode(education, 
             "1" = "Less than high school",
             "2" = "High school equivalent",
             "3" = "More than high school")) |> 
  drop_na() |> 
  filter(age >= 21)

nrow(demographic_clean)

accel_clean <- accel_raw |> 
  janitor::clean_names()

combined_accel <- left_join(demographic_clean, accel_clean, by = "seqn")
nrow(combined_accel)
head(combined_accel)
  
```



Produce a reader-friendly table for the number of men and women in each education category

```{r}
combined_accel |> 
  group_by(education, sex) |> 
  summarize(n_obs = n()) |> 
  pivot_wider(
    names_from = sex,
    values_from = n_obs,
  ) |> 
  knitr::kable(digits = 2)

```

create a visualization of the age distributions for men and women in each education category.

```{r}
combined_accel |> 
  ggplot(aes(x = age, y = sex)) +
  geom_density_ridges() +
  facet_grid(. ~ education)

combined_accel |> 
  ggplot(aes(x = age, y = sex)) +
  geom_density_ridges() +
  facet_grid(education ~ .)


```





